---
title: "MCM Project"
format: html
editor: visual
---

## Importing Libraries

```{r include=FALSE}
#Installing Packages
library(janitor)
library(Metrics)
library(readxl)
library(tidyverse)
library(tidymodels)
library("AmesHousing")
library(stringr)
library("olsrr")
library("glmnet")
library("doParallel")
library(readr)
library(maps)
library(viridis)
library(usmap)
library(kknn)
library(MASS)
library(rpart)
library(rpart.plot)
library(keras)
library(randomForest)
library(discrim)
library(caret)
library(gbm)
library(xgboost)
library(vip)
library(stringr)
library(AICcmodavg)
library(udpipe)
#all_cores<-parallel::detectCores()
#cl<- makePSOCKcluster(all_cores)
#registerDoParallel(cl)
```

## Importing Data

```{r}
data<-read_excel("Problem_C_Data_Wordle.xlsx", skip=1) %>% 
  clean_names()

#Removing words that arent
data<-data %>% 
  filter(word != "rprobe") %>% 
  filter(word != "clen") %>%
  filter(word != "tash")  %>% 
  filter(word != "marxh")
```

## Exploratory Analysis

```{r}
#Visualizing Relationships

#Comparing Date to number of people that participate
data %>% 
  ggplot(aes(x=date, y=number_of_reported_results))+
  geom_point()

#Are people getting better
data %>% 
  ggplot(aes(x=date, y=data$x7_or_more_tries_x))+
  geom_point()

#Are more people trying hard mode
data %>% 
  ggplot(aes(x=date, y=data$number_in_hard_mode))+
  geom_point()

#First attempt Guesses over time
data %>% 
  ggplot(aes(x=date, y=data$x1_try))+
  geom_point()



```

## Problem 1: Predicting number of players

This needs to be a regression model. This data is in no way linear so the first attempt is to create a nonlinear regression method.

```{r}
#Creating Data splits
set.seed(1234)

split<-initial_split(data, prop = .9)
training_data<-training(split)
test_data<-testing(split)

#The Future Date we want to predict the participants for
future_date<-data_frame(date = as.Date(x = "2023/3/1"),
                        contest_number = 620)

```

## Basic Decision Tree

```{r}
#Tuned Model
tuned_reg_tree<-decision_tree(mode = "regression", engine = "rpart", cost_complexity = 0.0005623413, tree_depth = 11, min_n = 2)

```

## Boosted Tree

```{r}
#Updating Workflow with Tuned Model and Fitting it to Training Data

tuned_reg_xgb<-boost_tree(mode = "regression", engine = "xgboost",mtry = 2, trees = 1000, min_n = 13, tree_depth = 6, learn_rate = 0.01872842, loss_reduction = 1.647096, sample_size = 0.9797893)

```

## Support Vector Machine Regression

```{r}
#Updating Workflow with Tuned Model and Fitting it to Training Data

tuned_svm<-svm_rbf(mode = "regression", engine = "kernlab", cost = 32, rbf_sigma = 1, margin = 0.1)

```

## Comparing Regressions

## Storing Tuned Models

```{r}
difficulty<-c()
for (i in 1:nrow(data)){
  current_row=data[i,]
  #Determining the difficulty of this word
  if (((current_row$x1_try + current_row$x2_tries + current_row$x3_tries) >
    (current_row$x3_tries + current_row$x4_tries)) && ((current_row$x1_try + current_row$x2_tries + current_row$x3_tries) >
    (current_row$x4_tries + current_row$x5_tries)) && ((current_row$x1_try + current_row$x2_tries + current_row$x3_tries) >
    (current_row$x5_tries + current_row$x6_tries)) && ((current_row$x1_try + current_row$x2_tries + current_row$x3_tries) >
    (current_row$x6_tries + current_row$x7_or_more_tries_x))){
    difficulty[i]<- "Very Easy"
  }else if (((current_row$x3_tries + current_row$x4_tries) >
    (current_row$x1_try + current_row$x2_tries + current_row$x3_tries)) && ((current_row$x3_tries + current_row$x4_tries) >
    (current_row$x4_tries + current_row$x5_tries)) && ((current_row$x3_tries + current_row$x4_tries) >
    (current_row$x5_tries + current_row$x6_tries)) && ((current_row$x3_tries + current_row$x4_tries) >
    (current_row$x6_tries + current_row$x7_or_more_tries_x))){
    difficulty[i]<- "Easy"
  }else if (((current_row$x4_tries + current_row$x5_tries) >
    (current_row$x1_try + current_row$x2_tries + current_row$x3_tries))== TRUE && ((current_row$x4_tries + current_row$x5_tries) >
    (current_row$x3_tries + current_row$x4_tries))== TRUE && ((current_row$x4_tries + current_row$x5_tries) >
    (current_row$x5_tries + current_row$x6_tries))== TRUE && ((current_row$x4_tries + current_row$x5_tries) >
    (current_row$x6_tries + current_row$x7_or_more_tries_x))== TRUE){
    difficulty[i]= "Medium"
  }else if (((current_row$x5_tries + current_row$x6_tries) >
    (current_row$x1_try + current_row$x2_tries + current_row$x3_tries)) && ((current_row$x5_tries + current_row$x5_tries) >
    (current_row$x3_tries + current_row$x4_tries)) && ((current_row$x5_tries + current_row$x6_tries) >
    (current_row$x4_tries + current_row$x5_tries)) && ((current_row$x5_tries + current_row$x6_tries) >
    (current_row$x6_tries + current_row$x7_or_more_tries_x))){
    difficulty[i]<- "Hard"
  }else if (((current_row$x6_tries + current_row$x7_or_more_tries_x) >
    (current_row$x1_try + current_row$x2_tries + current_row$x3_tries)) && ((current_row$x6_tries + current_row$x7_or_more_tries_x) >
    (current_row$x3_tries + current_row$x4_tries)) && ((current_row$x6_tries + current_row$x7_or_more_tries_x) >
    (current_row$x4_tries + current_row$x5_tries)) && ((current_row$x6_tries + current_row$x7_or_more_tries_x) >
    (current_row$x5_tries + current_row$x6_tries))){
    difficulty[i]<- "Very Hard"
  }else {difficulty[i]<-NA}
      
}
difficulty<-as_tibble(difficulty)

data$difficulty<-difficulty$value

```

```{r}
#Vowel Counting Function
get_vowel_count <- function(phrase) {
  counter <- 0
  for (i in unlist(strsplit(phrase, ""))) {
    if ( i %in% c("a", "e", "i", "o", "u")) {
      counter <- counter + 1 
    }   
  }
  counter
}

data %>% 
  filter(difficulty == "vh" | difficulty == "h")  
  

#Counting Vowels for every word
vowels<-c()
counter<-0
for (i in 1:nrow(data)){
  current_word = data$word[i]
  #print(current_word)
  vowels[i]<- get_vowel_count(current_word)
  #print(get_vowel_count(current_word))
}

data$vowel_count<-vowels

#Counting Number of consonants
data$consonant_count<-5-data$vowel_count

#Creating tricky letter function
get_tricky_letter_count <- function(phrase) {
  counter <- 0
  for (i in unlist(strsplit(phrase, ""))) {
    if ( i %in% c("r", "u", "w", "z", "x", "y")) {
      counter <- counter + 1 
    }   
  }
  counter
}

#Counting tricky letters
tricky_letters<-c()
for (i in 1:nrow(data)){
  current_word = data$word[i]
  #print(current_word)
  tricky_letters[i]<- get_tricky_letter_count(current_word)
  #print(get_tricky_letter_count(current_word))
}

data$tricky_letters<-tricky_letters

#Creating a function for counting letters
get_letter_count <- function(phrase) {
  letters<- as.vector(strsplit(phrase, ""))
  counter <- c()
  for (i in letters) {
 count <- str_count(phrase, i)
 counter[i]<-count
  }
  max(counter) 
}
#Counting the number of occurances of each letter
max_repeats<-c()
for (i in 1:nrow(data)){
  current_word = data$word[i]
  #print(current_word)
  max_repeats[i]<- get_letter_count(current_word)
  #print(get_tricky_letter_count(current_word))
}

data$max_repeats<-max_repeats
```

```{r}
#Checking which words dont have a difficulty
data$difficulty[data$word == "elder"]<- "Hard"
data$difficulty[data$word == "naïve" | data$word == "spell" | data$word == "berth"| data$word == "retro"| data$word == "girth"| data$word == "cater"| data$word == "mince"]<-"Medium"

#Making Difficulty a factor

data$difficulty<-as.factor(data$difficulty)
```

## EDA for Comparing Difficulties

```{r}
data %>% 
  group_by(difficulty) %>% 
  count
#Assess the vowel count over different difficulties
data %>% 
  ggplot(aes(x=difficulty, y=vowel_count))+
  geom_col(aes(fill = difficulty))

#Assessing Consonant Count Across Difficulties
data %>% 
  ggplot(aes(x=difficulty, y=consonant_count))+
  geom_col(aes(fill = difficulty))

#Assessing Tricky Letter Count Across Difficulties
data %>% 
  ggplot(aes(x=difficulty, y=tricky_letters))+
  geom_col(aes(fill = difficulty))

#Assessing Letter Repetition Across Difficulties
data %>% 
  ggplot(aes(x=difficulty, y=max_repeats))+
  geom_col(aes(fill = difficulty))

#Assessing Difficulty Over Time
data %>%
  ggplot(aes(x=date, y=difficulty))+
  geom_col(aes(fill = difficulty))

#Detecting parts of speech

library(udpipe)
library(dplyr)
udmodel <- udpipe_download_model(language = "english")
udmodel <- udpipe_load_model(file = udmodel$file_model)


wordle_pos <- udpipe_annotate(udmodel, 
                              data$word)
wordle_pos <- as.data.frame(wordle_pos)

wordle_pos %>% select(token, upos)

data$pos <- wordle_pos$upos




```

## Decision for Classifying Difficulty

```{r}
#Creating Data splits
set.seed(1234)

split<-initial_split(data, prop = .9)
training_data<-training(split)
test_data<-testing(split)

#The Future Date we want to predict the participants for
future_word<-data_frame(word = "eerie", vowel_count = 4, consonant_count =1, tricky_letters = 0, max_repeats = 3)

```

```{r}
#Decision Tree #
#tree_model<-decision_tree(mode = "classification",
#                          engine = "rpart",
#                          cost_complexity = tune(),
#                          tree_depth = tune(),
#                          min_n = tune())

#tree_wf<-workflow() %>% 
#  add_model(tree_model) %>% 
#  add_formula(difficulty ~ word + vowel_count + consonant_count + tricky_letters + max_repeats)

#folds<-vfold_cv(training_data)

#tree_tuning_grid<-grid_regular(cost_complexity(), tree_depth(), min_n(), levels = 5)

#tree_res<-tree_wf %>%
#  tune_grid(resamples = folds,
#            grid = tree_tuning_grid)

#best_tree_params<-tree_res %>% select_best("rmse")
```

```{r}
#Accessing model Accuracy
#tuned_tree<-finalize_model(tree_model, best_tree_params)
#tree_wf<-tree_wf %>% update_model(tuned_tree) %>% fit(training_data)

#tree_preds<-predict(tree_wf, test_data)

#tree_acc<-mean(tree_preds == test_data$difficulty)

#Predicting "eerie" difficulty
#predict(tree_wf, future_word)
```

## Boosted Tree Classification

```{r}
# Boosted Tree #

#xgb_classification<-boost_tree(mode = "classification", engine = "xgboost", mtry = tune(), trees = 1000, min_n = tune(), tree_depth = tune(), learn_rate = tune(), loss_reduction = tune(),sample_size = tune(), stop_iter = tune())

#xgb_wf<- workflow() %>% 
#  add_model(xgb_classification) %>% 
#  add_formula(difficulty ~ word + vowel_count + consonant_count + tricky_letters + max_repeats)

#folds<-vfold_cv(training_data)
#xgb_grid <- grid_latin_hypercube(
#  tree_depth(),
#  min_n(),
#  loss_reduction(),
#  sample_size = sample_prop(),
#  finalize(mtry(), training_data),
#  learn_rate(),
#  stop_iter(),
#  size = 30
#)

#xgb_res<- xgb_wf %>% 
#  tune_grid(resample = folds, grid = xgb_grid, control = control_grid(save_pred = TRUE))

#best_xgb_params<- select_best(xgb_res, "accuracy")
```

```{r}
  
```
